\normalfont \LARGE Improved Quasi-Static Methods for Time-Dependent Neutron Diffusion and Implementation in Rattlesnake

### 2016-10-13 doc/Thesis_proposal/thesis_proposal.tex
Thesis: Addressing some comments

Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature.  However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.  Transient computational methods currently in \tcr{laboratory frameworks--people may not understand what you mean here} have proven to be overbearingly \tcr{slow-CPU-expensive} for complex, real-world scenarios.  Therefore, more efficient and faster methods for computational \tcr{frameworks} are highly desired.  The Department of Energy (DOE) has especially sought to develop in the MOOSE framework.  \\

@@

The difficulty in transient reactor modeling is due to the neutron transport equation and \tcr{the complicated nature of neutron motion that the equation represents-- ???? it is the dimensionality of the phase-space that is the issue, and the fact that the time discretization has to be implicit, which  leads to kinear system solves}.  The equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.  Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.  For the purposes of this research, neutron diffusion will be discussed exclusively.  \tcr{Additionally, simulations can often neglect temporal dependence due to the steady-state nature of reactors--???}.  However, simulations for facilities such as TREAT, must retain time dependence in order to obtain relevant information. \\

For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.  There are many discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research \tcr{I do not know of ``many'' discretizations in the energy variable ...}.  Since the purpose of this research involves temporal dependence, discretization in time was kept general.  The time dependent neutron diffusion equation with delayed neutron precursors can be seen in Equations~\ref{eq:flux}~and~\ref{eq:precursor}.

@@

\tcr{is it interesting to try to converge the nonlinearities in $T$ and $\rho$ before going back to the shape?}

@@

A one-dimensional, single group FEM prototype will be created for IQS in order to test validation for simple problems. However, testing complex, real-world problems is most easily performed by implementing IQS in a FEM framework.  The Rattlesnake radiation transport framework is a excellent option for performing these problems.  Rattlesnake already contains a neutron transport method for CFEM diffusion equivalent to the brute-force method \tcr{Are you referring to SAAF? or did you mean diffusion in lieu of transport?}.  Additionally, Rattlesnake is well documented and well organized, so implementing IQS may be relatively straightforward. With this implementation, benchmark testing such as 16-A1 \cite{ANL_BPB}, TWIGL, LRA, C5G7-2D, etc. can be performed, as well as a full core simulation of the TREAT reactor.

>>>>

Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature.  However, recent developments in nuclear testing, including the revitalization of the Transient Reactor Testing (TREAT) Facility at Idaho National laboratory (INL), have brought significant attention to the development of transient reactor modeling.  Transient computational methods currently in \tcr{laboratory frameworks--people may not understand what you mean here} have proven to be overbearingly computationally expensive for complex, real-world scenarios.  Therefore, more efficient and faster methods for computational \tcr{frameworks} are highly desired.  The Department of Energy (DOE) has especially sought to develop in the MOOSE framework.  \\

@@

The difficulty in transient reactor modeling is due to the neutron transport equation, which describes the flux of neutrons. The difficulty in solving or evaluating the equation comes from the dimensionality of the phase-space and the fact that the time discretization has to be implicit, which  leads to kinear system solves.  The equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.  Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.  For the purposes of this research, neutron diffusion will be discussed exclusively.  \tcr{Additionally, simulations can often neglect temporal dependence due to the steady-state nature of reactors--???}.  However, simulations for facilities such as TREAT, must retain time dependence in order to obtain relevant information. \\

For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.  There are several viable discretization schemes for each of these variables, but FEM for space and multigroup method for energy were used for the development of this research. Since the purpose of this research involves temporal dependence, discretization in time was kept general.  The time dependent neutron diffusion equation with delayed neutron precursors can be seen in Equations~\ref{eq:flux}~and~\ref{eq:precursor}.

@@

The couplings between temperature, amplitude, and shape are all nonlinear, so iteration processes are need for each time scale.  The amplitude and temperature need to be iterated on the middle time scale until convergence on each temperature step.  Then another iterative process needs to occur in the shape time scale scale on all three variables.

@@

A one-dimensional, single group FEM prototype will be created for IQS in order to test validation for simple problems. However, testing complex, real-world problems is most easily performed by implementing IQS in a FEM framework.  The Rattlesnake radiation transport framework is a excellent option for performing these problems.  Rattlesnake already contains a neutron diffusion method for continuous-FEM equivalent to the brute-force method.  Additionally, Rattlesnake is well documented and well organized, so implementing IQS may be relatively straightforward. With this implementation, benchmark testing such as 16-A1 \cite{ANL_BPB}, TWIGL, LRA, C5G7-2D, etc. can be performed, as well as a full core simulation of the TREAT reactor.
